### 2014-01-27 brains-protocol.tex
3.4 is reread and good

\subsubsection{Network-based indices}

@@

\begin{figure}[!th]
    \includegraphics[width=\linewidth]{algorithms/Rplot.pdf}
\end{figure}

@@

We then measure the cognitive alteration of human copy-pasting by comparing the relative properties of substituted and substituting words in quotations.  We only consider altered quotations, \hbox{i.e.} quotations where at least one word has been modified; in other words, we measure the properties of alteration \emph{knowing that} there has been a variation (which indicates a certain reformulation) while we do not take invariant quotations into account, as it is impossible to know, in that case, whether there has been perfect individual reformulation, or just digital copy-pasting of a source (``{\sf Ctrl-C}/{\sf Ctrl-V}'').

We build two main observables for each word property. First, we measure the variation of word features over a substitution, looking at the difference of the value of a given feature between arrival and start words. %This address the question ``how do word features vary over a subs  \tb{la variation de caractéristique sémantique lors du passage du mot remplacé au mot remplaçant dans une substitution. Ceci répond à la question "comment fait-on varier la caractéristique sémantique lorsqu'on substitue un mot ?".} 
Second, we measure the susceptibility for words to be the target of a substitution, knowing that there has been a variation, in order to show which semantic features are the most likely to be the target of a substitution. % Second, \tb{la susceptibilité que les mots ont à être remplacés, en fonction de leur caractéristique sémantique. Cette mesure répond à la question "quelles caractéristiques sémantiques a-t-on plus tendance à substituer ?"%\footnote{Attention, cette mesure ne suffit pas en elle-même pour parler d'\emph{influence} de la caractéristique sémantique d'un mot sur le fait qu'il soit substitué ou non, parce qu'on n'a pas fait d'analyse des causes impliquées dans le fait qu'un mot soit substitué ; en effet, d'autres facteurs pourraient être plus importants que ces caractéristiques sémantiques dans les causes d'une substitution. On ne répond donc \emph{pas} à la question "quelle est l'influence de la caractéristique sémantique d'un mot sur sa propension à être substitué ?".}.

\subsubsection{Alteration}
%The above protocol enables us to observe in detail how word features vary upon substitution. 
For a given feature $\mathfrak{F}$, we measure how a word $\wstart$'s feature varies as $\wstart$ is substituted by $\warrival$. Let us denote this quantity:
$$\Delta (\wstart,\warrival) \defeq \mathfrak{F}(\warrival) - \mathfrak{F}(\wstart)$$

Averaging this value over all start words with a given feature value $f$ yields the mean variation for that feature value $f$.\footnote{To avoid any auto-correlation effect due to the number of substitutions in a cluster, we first average substitutions over each cluster, by considering the average of arrival word features for a given starting word. Indeed, substitutions occurring in the same cluster are likely not statistically independent.} This 
quantity can be written:
$$\left< \Delta(\wstart,\warrival) \right>_f = \left< \mathfrak{F}(\warrival) - \mathfrak{F}(\wstart) \right>_{\left\lbrace (\wstart,\warrival) | \mathfrak{F}(\wstart) = f \right\rbrace}$$

We introduce a null hypothesis $\mathcal{H}_0$ to compare the actual variation of a word's feature to its expected variation assuming the arrival word $\warrival_{\mathcal{H}_0}$ had been chosen randomly from % , let us denote the corresponding quantity This value is to be compared to the variation the word's feature would have undergone had the arrival word been chosen randomly in 
the pool of \TB{WN/FA} words. The corresponding average quantity over all start words may be written as:\footnote{Note that $\warrival_{\mathcal{H}_0}$ is in fact constant in this averaging, since by definition it does not depend on $\wstart$.}
%$$\Delta_{\mathcal{H}_0} (\wstart,\warrival) \defeq f\left(\warrival_{\mathcal{H}_0}\right) - f\left(\wstart\right)$$
%and, similarly, we compute the average over all start words 
$$\left< \Delta(\wstart,\warrival_{\mathcal{H}_0}) \right>_f
= \left< \mathfrak{F}({\warrival_{\mathcal{H}_0}}) - \mathfrak{F}(\wstart) \right>_{\left\lbrace (\wstart,\warrival_{\mathcal{H}_0}) | \mathfrak{F}(\wstart) = f \right\rbrace}$$
%Averaging over all start words with a given feature value $f$, we get $$\left< \Delta_{\mathcal{H}_0} \right>_f = \left< \mathfrak{F}({\warrival_{\mathcal{H}_0}}) - \mathfrak{F}(\wstart) \right>_{\left\lbrace \wstart | \mathfrak{F}(\wstart) = f \right\rbrace}$$

We also considered an alternative null hypothesis, denoted $\mathcal{H}_{00}$, where the arrival word is chosen randomly \emph{among immediate synonyms of the start word} (neighbors), i.e. an arrival word chosen among semantically plausible though still random words (in this case $\warrival_{\mathcal{H}_{00}}$ does depend on $\wstart$). %The values for $d$ were decided upon according to the distribution of distances travelled on the synonym graph upon substitution, normalized by the number of synonyms at that distance. 
%  brought all variation curves shown below nearer to the origin, but did not change the results qualitatively.
%The variation of feature compared to $\mathcal{H}_0$ is then $\left< \Delta \right>_f - \left< \Delta_{\mathcal{H}_0} \right>_f$.

@@

This measure focuses on the selection of start word instead of the selection of the arrival word. Indeed, the features have an effect not only on the choice of a new word when a substitution takes place, but also at the preceding moment when it is not yet known which word in the quote --~if any~-- will be substituted.

>>>>

\subsubsection{Network-based measures}

@@

\begin{figure}[!th]
    \includegraphics[width=\linewidth]{algorithms/Rplot.pdf}
    \caption{Variable correlations \rk{rebuild with the variables we talk about and keep}}
\end{figure}

@@

We then measure the alterations introduced when authors reproduce quotations by comparing the relative features of substituted and substituting words.
Note that since we only consider substitutions, we measure the features of an alteration \emph{knowing that there has been a alteration}, and we do not take invariant quotations into account.
Indeed, in the first case we know there has been a human reformulation, whereas in the second case it is impossible to know whether there has been perfect human reformulation or simply digital copy-pasting of a source (``{\sc Ctrl-C}/{\sc Ctrl-V}'').

We build two main observables for each word feature.
First, we measure the variation of word feature over a substitution, looking at the variation of a given feature between arrival and start words.
Second, we measure the susceptibility for words to be the source of a substitution, knowing that there has been a variation, in order to show which semantic features are the most likely to attract a substitution.

\subsubsection{Alteration}

For a given feature $\mathfrak{F}$, we measure how a word $\wstart$'s feature varies as $\wstart$ is substituted by $\warrival$.
Let us denote this quantity:
$$\Delta (\wstart,\warrival) \defeq \mathfrak{F}(\warrival) - \mathfrak{F}(\wstart)$$

Averaging this value over all start words with a given feature value $f$ yields the mean variation for that feature value $f$.
\footnote{To avoid any auto-correlation effect due to the number of substitutions in a cluster (possibly leading to an overly optimistic estimation of confidence intervals), we first average substitutions over each cluster, by considering the average of arrival word features for a given start word.
Indeed, substitutions occurring in the same cluster are likely not statistically independent.}
This quantity can be written:
$$\left< \Delta(\wstart,\warrival) \right>_f = \left< \mathfrak{F}(\warrival) - \mathfrak{F}(\wstart) \right>_{\left\lbrace (\wstart,\warrival) | \mathfrak{F}(\wstart) = f \right\rbrace}$$

We introduce a null hypothesis $\mathcal{H}_0$ to compare the actual variation of a word's feature to its expected variation, assuming the arrival word $\warrival_{\mathcal{H}_0}$ had been chosen randomly from the pool of free association words.
\rk{Check that's true in the code}
The corresponding average quantity over all start words may be written:
\footnote{Note that $\warrival_{\mathcal{H}_0}$ is in fact constant in this averaging, since by definition it does not depend on $\wstart$.}
$$\left< \Delta(\wstart,\warrival_{\mathcal{H}_0}) \right>_f = \left< \mathfrak{F}({\warrival_{\mathcal{H}_0}}) - \mathfrak{F}(\wstart) \right>_{\left\lbrace (\wstart,\warrival_{\mathcal{H}_0}) | \mathfrak{F}(\wstart) = f \right\rbrace}$$

We also considered an alternative null hypothesis, denoted $\mathcal{H}_{00}$, where the arrival word is chosen randomly \emph{among immediate synonyms of the start word} (neighbours in the WN network \rk{We could also do that with neighbours in directed or undirected FA}), i.e. an arrival word chosen among semantically plausible though still random words.
\footnote{In this case $\warrival_{\mathcal{H}_{00}}$ does depend on $\wstart$.}

@@

This measure focuses on the selection of start word instead of the selection of the arrival word.
Indeed, the features have an effect not only on the choice of a new word when a substitution takes place, but also at the preceding moment when it is not yet known which word in the quotation -- if any -- will be substituted.
